A formal approach but practical approach in train regulation

### 2025-06-12 paper/main.tex
regulator 1

\usepackage{package/regles}
\usepackage{mathtools}

@@

@@

\newcommand{\sigred}{{\mbox{$\mathtt{r}$}}\xspace}
\newcommand{\siggreen}{{\mbox{$\mathtt{g}$}}\xspace}

\newcommand{\deviate}{{\mbox{$\mathtt{v}$}}\xspace}
\newcommand{\direct}{{\mbox{$\mathtt{d}$}}\xspace}

@@

Now, we come to the train model. Each train has an identifier. In other words, we have a set of trains $\trains = \{1, \ldots, k\}$. Each train has a position on the network. The position is the index of the block it stays in or the last block it was in. Next, a train has a direction in $\directions$. The direction does not mean the train is running---it can be blocked by some red signal---but that its next move will be in that direction. Finally, a train has orders. Orders are given as a sequence of orders following the grammar:
$$  seq ::= \su(dir,pos) $$
with $dir \in \directions$ and $pos \in \blocks$. An order $\su(dir, pos)$ states that the train direction is set to $dir$ and that the train runs until it reaches $pos$.  The sequence may be empty.

@@

The regulator is characterized as follows. First, it has a general program, next mentioned as $E$. The program contains the plan of the regulator. The regulator will follow its rule. A program is given as a sequence of statements. Those follow the grammar:

\begin{grammar}[Regulator's rules][h][regulateur]
    \firstcasesubtil{stmt}{ev(id, n) : atoms}{}
    \firstcasesubtil{atoms}{incr(j)}{}
    \otherform{att(j, value_{j})}{}
    \otherform{turn(id_{switch}, state)}{}
    \otherform{auth}{}
    \otherform{atoms \gralt \varepsilon}{}
\end{grammar}

>>>>

\usepackage{package/regles}
\usepackage{mathtools}
\usepackage{ amssymb }

@@

\newcommand{\Nat}{{\mbox{$\mathbb{N}$}}}

@@

%% tt style
\newcommand{\sigred}{{\mbox{$\mathtt{r}$}}\xspace}
\newcommand{\siggreen}{{\mbox{$\mathtt{g}$}}\xspace}

\newcommand{\deviate}{{\mbox{$\mathtt{v}$}}\xspace}
\newcommand{\direct}{{\mbox{$\mathtt{d}$}}\xspace}
\newcommand{\incr}{{\mbox{$\mathtt{incr}$}}\xspace}
\newcommand{\turn}{{\mbox{$\mathtt{turn}$}}\xspace}

@@

Now, we come to the train model. Each train has an identifier. In sidother words, we have a set of trains $\trains = \{1, \ldots, k\}$. Each train has a position on the network. The position is the index of the block it stays in or the last block it was in. Next, a train has a direction in $\directions$. The direction does not mean the train is running---it can be blocked by some red signal---but that its next move will be in that direction. Finally, a train has orders. Orders are given as a sequence of orders following the grammar:
$$  seq ::= \su(dir,pos) $$
with $dir \in \directions$ and $pos \in \blocks$. An order $\su(dir, pos)$ states that the train direction is set to $dir$ and that the train runs until it reaches $pos$.  The sequence may be empty.

@@

The regulator is characterized by 6 fields. First, it has a general program, next mentioned as $E$. The program contains the plan of the regulator. The regulator will follow its rule. A program is given as a sequence of statements. Those follow the grammar:

\begin{grammar}[Regulator's rules][h][regulateur]
    \firstcasesubtil{stmt}{ev(id, n) : atoms}{}
    \firstcasesubtil{atoms}{\incr(j)}{}
    \otherform{att(j, v)}{}
    \otherform{\turn(sid, s)}{}
    \otherform{auth}{}
    \otherform{atoms ; atoms}{}
    \otherform{\varepsilon}{}
\end{grammar}
with $id \in \trains$, $n \in  \Nat$, $j \in \blocks$, $sid \in \turnouts$, $s\in \internalState$.  All rules are guarded by a condition $ev(id,n)$ stating that train $id$ reached the $n$-th step of its route. The atoms update the regulator fields. Before we describe them and the other fields, a notation: given some function $f: X \to Y$, $x \in X$ and $y \in Y$, we denote by $f[x \leftarrow y]$, the function:
 $$f[x \leftarrow y] \triangleq \left\{  \begin{array}{ll} x &\mapsto y\\
x' \neq x \in X &\mapsto f(x')
\end{array}\right.$$

The token field of the regulator, denoted $T$, is a function  $T: \blocks \to \Nat$ associating a token level to each block.  The atom $\incr(j)$ updates the token $j$:  $T \cdot \incr(j)  \triangleq T[j \leftarrow T(j)+1]$.

The switch field is a function $S: \turnouts \to \internalState$ giving the network configuration. Again, the atom $\turn(sid,s)$ updates the internal state of $id$: $S \cdot \turn(sid,s) \triangleq S[sid \leftarrow s]$.

The waiting field is a function $W: \blocks \times \Nat \to \trains \times Bool$. XXX

The buffer field is the XXX

The history field $H: \trains \to \blocks$ gives the last position of each train recorded  by the regulator.

So, the regulator is described as a tuple $R = (E, T, W, B, H)$. Applying $\incr(j)$ to the regulator is applying it to the token field: $R \cdot \incr(j) = (E, T \cdot \incr(j), W, B, H)$. We use the same convention for the other atoms.
